Grade Retention and Mixed Race

### 2022-08-31 products/paper/main.qmd
Restructure headings; correct description of calculating child race

# Background

## Where do Biracials Fit?

@@

Second, even when "correctly" classified, greater physical ambiguity may allow biracial individuals to avoid the worst instances of bias and discrimination because they do not conform to the ideal type for the group facing the most discrimination. In other words, the experience of discrimination may be "softened" because their ambigious physical appearance provokes less racial antipathy. The US in fact already has a long history of this process. Because of the one-drop rule, Black individuals vary substantially in skin tone and those individuals with a lighter skin tone have better outcomes. Thus light skin can moderate (but not eliminate) the stigma of Blackness even among monoracial individuals.

@@

## Prior Work

@@

## Grade Retention as an Outcome

@@

The race of children is calculated by a cross-tabulation of parent's races. A full cross-tabulation of the race/Hispanicity of each parent would produce far too many combined categories to be analytically useful. Therefore, I collapse each parent's race into the categories of White, Black, Asian, Latino, and Indigenous. These categories form the "ethnoracial pentagon" that is commonly used in popular practice and government tabulation to identify race in the US [@hollinger_postethnic_1995]. Respondents are identified as Latino base on their response to the Hispanicity question, regardless of their response to the race question. Respondents are identified as Indigenous if they responded as either American Indian/Alaska Native or as Pacific Islander. Because the goal of my analysis is to examine specifically the outcomes of "first-generation" biracial children, I exclude cases where at least one parent identified with multiple races.

>>>>

# Where do Biracials Fit?

@@

Second, even when "correctly" classified, greater physical ambiguity may allow biracial individuals to avoid the worst instances of bias and discrimination because they do not conform to the ideal type for the group facing the most discrimination. In other words, the experience of discrimination may be "softened" because their ambiguous physical appearance provokes less racial antipathy. The US in fact already has a long history of this process. Because of the one-drop rule, Black individuals vary substantially in skin tone and those individuals with a lighter skin tone have better outcomes. Thus light skin can moderate (but not eliminate) the stigma of Blackness even among monoracial individuals.

@@

# Prior Work

@@

# Grade Retention as an Outcome

@@

The race of children is calculated by a cross-tabulation of parent's races. A full cross-tabulation of the race/Hispanicity of each parent would produce far too many combined categories to be analytically useful. Therefore, I collapse each parent's race into the categories of White, Black, Asian, Latino, and Indigenous. These categories form the "ethnoracial pentagon" that is commonly used in popular practice and government tabulation to identify race in the US [@hollinger_postethnic_1995]. Parents are identified as Latino based on their response to the Hispanicity question, regardless of their response to the race question. Parents are identified as Indigenous if they responded as either American Indian/Alaska Native or as Pacific Islander. Because the goal of my analysis is to examine specifically the outcomes of "first-generation" biracial children, I exclude cases where at least one parent identified with multiple races.
